---
title: "p8105_hw2_cl4044"
author: "Chenxi Liu"
date: "9/26/2020"
output: github_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1

First, define a path to the dataset. 

```{r}
path_to_data = "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx"
```


Read the Mr. Trashwheel dataset. 

```{r}
trashwheel_df = 
	read_xlsx(
		path = path_to_data,
		sheet = "Mr. Trash Wheel",
		range = cell_cols("A:N")) %>% 
	janitor::clean_names() %>% 
	drop_na(dumpster) %>% 
	mutate(
		sports_balls = round(sports_balls),
		sports_balls = as.integer(sports_balls)
	)
```

Read precipitation data! For 2018 and 2017. 

```{r}
precip_2018 = 
	read_excel(
		"./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "2018 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2018) %>% 
	relocate(year)
precip_2017 = 
	read_excel(
		"./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "2017 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2017) %>% 
	relocate(year)
```

Now combine annual precipitation dataframes.

```{r}
month_df = 
	tibble(
		month = 1:12,
		month_name = month.name
	)
precip_df = 
	bind_rows(precip_2018, precip_2017)
precip_df =
	left_join(precip_df, month_df, by = "month")
precip_df
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data. In this dataset:

* The median number of sports balls found in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`
* The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.

## Problem 2

Read and clean the NYC Transit data. 

```{r, message = FALSE}
transit_df = 
	read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
	janitor::clean_names() %>% 
  select(-division, -staffing, -staff_hours, -ada_notes, -free_crossover, -north_south_street, -east_west_street, -corner, -entrance_latitude, -entrance_longitude, -station_location, -entrance_location, -exit_only ) %>%
  mutate(entry = recode(entry, "YES" = "TRUE", "NO" = "FALSE"),
         vending = recode(vending, "YES" = "TRUE", "NO" = "FALSE"),
         entry = as.logical(entry),
         vending = as.logical(vending)
  )
transit_df
str(transit_df)
```
This dataset contains  – explain briefly what variables the dataset contains, describe your data cleaning steps so far, and give the dimension (rows x columns) of the resulting dataset. Are these data tidy?

